---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(data.table)
library(caret)
library(reticulate)
library(tensorflow)
library(keras)


# df<-read_rds("/Users/shivangi/Fundamentals-of-AI/yg821jf8611_ma_statewide_2020_04_01.rds")
df<-fread("/Users/shivangi/fundamentals_of_ai/ma_statewide_2020_04_01.csv" )

head(df,10)

```

#### Data preprocessing

Removing location, raw_row_number, because they have high cardinality, and serve no purpose in our analysis. 

```{r}

df<-select(df,-c("location","raw_row_number","arrest_made","citation_issued","warning_issued","type","raw_Race"))

df$date<-as.Date(df$date)

df<-df%>%filter((subject_race%in%c("unknown","other"))==FALSE)

df<-df%>%mutate(year=year(date))%>%
         mutate(month=month(date))%>%
         mutate(weekday=weekdays(date))%>%
         mutate(dayofmonth=as.numeric(format(df$date,'%d')))


df<-df%>%select(-c("date"))

df$county_name<-df$county_name%>%replace_na("unknown")
df<-df[!is.na(df$subject_race), ]
df<-df[!is.na(df$subject_sex), ]
df<-df[!is.na(df$outcome), ]


df$search_basis<-df$search_basis%>%replace_na("unknown")
df<-df%>%mutate_if(is.logical,as.numeric)
df<-df%>%replace(is.na(.),99)


```


```{r}

# label<-as.matrix(df$subject_race)
# df<-df%>%select(-subject_race)

dummy <- dummyVars("~ .", data=df)
newdata <- data.frame(predict(dummy, newdata = df))

rm(df)
set.seed(123)

#splitting data into train and test
train_index <- sample(seq_len(nrow(newdata)),size = floor(0.70*nrow(newdata)))


train<-newdata[train_index,]
test<-newdata[-train_index,]

# train_label<-label[train_index,]
# test_label<-label[train_index,]

# y_train<-as.numeric(factor(train_label))%>%
#          matrix(nrow = 2364527, ncol = 1) %>% 
#         to_categorical(num_classes = 0:3)

y_train<-cbind(train$subject_raceasian.pacific.islander,train$subject_raceblack,train$subject_racehispanic,train$subject_racewhite)%>%matrix(nrow = 2364527, ncol = 4)

y_test<-cbind(test$subject_raceasian.pacific.islander,test$subject_raceblack,test$subject_racehispanic,test$subject_racewhite)%>%matrix(nrow=1013369,ncol=4)

train<-train%>%select(-subject_raceasian.pacific.islander,-subject_raceblack,-subject_racehispanic,-subject_racewhite)
#%>% matrix(nrow = 2364527, ncol = 108)

test<-test%>%select(-subject_raceasian.pacific.islander,-subject_raceblack,-subject_racehispanic,-subject_racewhite)

#%>%matrix(nrow=1013369,ncol=108)
train<-drop_na(train)

```



```{r}
# create model
model <- keras_model_sequential()
# define and compile the model
model %>% 
  layer_dense(units = 100, activation = 'relu', input_shape = c(108)) %>% 
  layer_dropout(rate = 0.5) %>% 
  layer_dense(units = 64, activation = 'relu') %>% 
  layer_dropout(rate = 0.5) %>% 
  layer_dense(units = ncol(y_train), activation = 'softmax') %>% 
  compile(
    loss = 'categorical_crossentropy',
    optimizer = optimizer_sgd(lr = 0.01, decay = 1e-6, momentum = 0.9, nesterov = TRUE),
    metrics = c('accuracy')     
  )
# train
model %>% fit(train, y_train, epochs = 20, batch_size = 128)

```



```{r}
summary(model)

```


```{r}

#generate dummy data
x_train <- matrix(runif(10000*108), nrow = 10000, ncol = 108)
y_train <- runif(10000, min = 0, max = 3) %>%
  round() %>%
  matrix(nrow = 10000, ncol = 1) %>%
  to_categorical(num_classes = 4)
x_test  <- matrix(runif(5000*108), nrow = 5000, ncol = 108)
y_test <- runif(5000, min = 0, max = 3) %>%
  round() %>%
  matrix(nrow = 5000, ncol = 1) %>%
  to_categorical(num_classes = 4)

```




```{r}

# # create model
# model <- keras_model_sequential()
# # define and compile the model
# model %>% 
#   layer_dense(units = 64, activation = 'relu', input_shape = c(108)) %>% 
#   layer_dropout(rate = 0.5) %>% 
#   layer_dense(units = 64, activation = 'relu') %>% 
#   layer_dropout(rate = 0.5) %>% 
#   layer_dense(units = 4, activation = 'softmax') %>% 
#   compile(
#     loss = 'categorical_crossentropy',
#     optimizer = optimizer_sgd(lr = 0.01, decay = 1e-6, momentum = 0.9, nesterov = TRUE),
#     metrics = c('accuracy')     
#   )
# # train
# model %>% fit(x_train, y_train, epochs = 20, batch_size = 128)


```



```{r}

# fit model with our training data set, training will be done for 100 times data set
# Fit the model 

fitted_model<-fit(model,
                  train, 
                  as.matrix(y_train), 
                  epochs = 30, 
                  batch_size=512,
                  validation_split = 0.3,
                  verbose=0
                  )

#Visualize trained model
plot(fitted_model)
```





### Neural network evaluation

After training, we evaluate the model using evaluate(). The result shows us the loss and mean absolute error, which is 0.05. We have hence acheived our prescribed accuracy. We generate the predictions using the predict() function. Visualizing the performance, we see that the model approximates the function decently well. 

```{r}
#Evaluate model
model%>%evaluate(test,y_test)

#Produce predicted y
# pred<- model %>% predict(test %>% select(-y))



```